\texttt{ripple}: Differentiable Waveforms for Gravitational Wave Data Analysis

### 2022-09-29 src/tex/ms.tex
Added first figure and nearly finised section 2

There are a variety of gravitational waveforms currently used in analysis pipelines.
They are generally structured into different families, the most common of which are: the effective one body, the phenomenological inspiral-merger-ringerdown (IMR), and numerical relativity surrogate.
Of these, the phenomenological IMR family is most well suited for an AD implementation in JAX.
In particular, its closed form expression, ... \TE{Why else?}
\te{Do we need more of an intro for why IMRPhenom?}

@@

To ensure our implempation of IMRPhenomD is faithful to the lalsuite implementation, we start by defining the noise weighted inner product:
\begin{equation}
    \left(h_{\theta_1}|h_{\theta_2}\right) \equiv 4 \, \mathrm{Re} \int^{\infty}_{0} \mathrm{d} f \, \frac{ h_{\theta_1}(f) h^*_{\theta_2}(f)}{S_n(f)}\, ,
\end{equation}
where $S_n$ is the (one-sided) noise power spectral density (PSD) and $h_{\theta_1}$ is the frequency domain waveform evaluated at the instrinsic parameter point $\theta_1$.
We can then normalize the inner product through
\begin{equation}
    \left[h_{\theta_1}|h_{\theta_2}\right] = \frac{\left(h_{\theta_1}|h_{\theta_2}\right)}{\sqrt{\left(h_{\theta_1}|h_{\theta_1}\right)\left(h_{\theta_2}|h_{\theta_2}\right)}}\, .
\end{equation}
Now we are ready to define the \textit{match}

\begin{figure}[ht!]
    \script{random_numbers.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/random_numbers.pdf}
        \caption{
            This plot should show a match colour bar.
        }
        \label{fig:random_numbers}
    \end{centering}
\end{figure}

\subsection{Benchmarks}
\label{subsec:benchmarks}

@@

\subsection{Fisher Analysis}
\label{subsec:fisher}

>>>>

There are a variety of gravitational waveforms currently used in analysis pipelines.
They are generally structured into different families, the most common of which are: the effective one body, the phenomenological inspiral-merger-ringerdown (IMR), and numerical relativity surrogate.
Of these, the phenomenological IMR family is most well suited for an AD implementation in \texttt{jax}.
In particular, its closed form expression, ... \TE{Why else?}
\te{Do we need more of an intro for why IMRPhenom?}

@@

To ensure our implempation of IMRPhenomD is faithful to the lalsuite implementation, we start by defining the noise weighted inner product:
\begin{equation}
    \left(h_{\theta_1}|h_{\theta_2}\right) \equiv 4 \, \mathrm{Re} \int^{\infty}_{0} \mathrm{d} f \, \frac{ h_{\theta_1}(f) h^*_{\theta_2}(f)}{S_n(f)}\, ,
\end{equation}
where $S_n$ is the (one-sided) noise power spectral density (PSD) and $h_{\theta_1}$ is the frequency domain waveform evaluated at the instrinsic parameter point $\theta_1$.
We can then normalize the inner product through
\begin{equation}
    \left[h_{\theta_1}|h_{\theta_2}\right] = \frac{\left(h_{\theta_1}|h_{\theta_2}\right)}{\sqrt{\left(h_{\theta_1}|h_{\theta_1}\right)\left(h_{\theta_2}|h_{\theta_2}\right)}}\, .
\end{equation}
Now we are ready to define the \textit{match} which is given by 
\begin{equation}
    \mathrm{m}(h_{\theta_1}|h_{\theta_2}) \equiv \max_{t_c,\, \phi_c} \left[h_{\theta_1} \, \middle| \, h_{\theta_2} \right]\,
\end{equation}
where $t_c$ and $\phi_c$ are, respectively, the time and phase of coalesence.

Since the match is a measure of the difference between two waveforms, we can use it to demonstrate that the implementation of IMRPhenomD in \texttt{ripple} accurately matches the \texttt{lalsuite} implementation.
This is shown in Fig.~\ref{fig:match}, where we have calculated the $\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})$ across the entire parameter space.\footnote{
    Specifically, we use 2000 points in the ranges $m_{1,2} = (1,50)\,M\_{\odot}$ and $\chi_{1,2} = (-1,1)$. In addition, we evaulated the waveforms on a frequency grid from $32\,$Hz to $1024\,$Hz with frequency spacing $\Delta f = 0.0125 \,$Hz.
}
Here, $h_{\theta_1}$ corresponds to the \texttt{ripple} waveform implementation and $h^{*}_{\theta_1}$ is the \texttt{lalsuite} implementation. 
From Fig.~\ref{fig:match}, it is clear that the \texttt{ripple} waveform matches with the \texttt{lalsuite} waveform to numerical precision across the majority of the parameter space.
In particular, the blue points indicate parameter points where the $1-\mathrm{m}(h_{\theta_1}|h^{*}_{\theta_1})=0$ and would therefore not appear in the log scale colorbar.
\begin{figure}[ht!]
    \script{random_matches.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/random_matches.pdf}
        \caption{
            This plot should show a match colour bar.
        }
        \label{fig:match}
    \end{centering}
\end{figure}

There remains some slight deviation between the two waveform implementations at higher masses. 
This is likely due to the fact that cubic interpolators, which are used within IMRPhenomD to calculate the ringdown and damping frequencies, are not currently supported in \texttt{jax}.
Instead, we initially use \texttt{scipy}'s cubic interpolator to create a fine grid of XXX values, which we then linearly interpret during waveform generation.
Unfortunately, we cannot make the initial cubic interpolation arbitrarily fine as this would add additional computational overhead when loading the data during waveform evaluation.
The remaining error in the waveform at high masses is therefore primarily down to the coarseness in this interpolation.
Note however, that the differences are well below the accuracy requirements of the waveforms and will have no impact on any data analysis tasks.

In order to be usuable in most data analysis tasks, a waveform needs to be fast to evaluate.
Fortunately, the Phenom IMR waveforms are constructed from simple closed form expressions which are therefore computationally efficient.
The primary arguement of this paper is that waveform derivatives will also be invauable to data analysis tasks.

@@
